LAs_clusters

### 2020-09-11 src/LAs_Clusters.Rmd
the N in the regression table is not correct. it should be equal to the min N from the descriptive stats.

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(rprojroot)
library(lubridate)
library(knitr)
library(rprojroot)
library(nnet)
library(stargazer)
library(openair)
library(rgdal)
library(corrplot)
library(DescTools)
library(geosphere)

@@

# LAD codes and names
# get LA 
# this is the heaviest version:
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
# path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
# la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC: 
la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
codes <- la@data %>%
  select(lad19cd, lad19nm) %>%
  rename(LAD19NM = lad19nm)

@@

ladDist$distMet <- apply(ladDist[,3:14], 1, min)

@@

## Job density

```{r}
job.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_57_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318&date=latest&item=3&measures=20100")

@@

## Population

@@

## Labour supply

```{r}
labour <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_17_5.data.csv?geography=1816133633...1816133837,1820327937...1820328307&date=latest&variable=18,1532...1540,290,720&measures=20599,21001,21002,21003")

@@

## Earnings

```{r}
earnings <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_30_1.data.csv?geography=1946157348,2092957698,2013265929,1816133633...1816133837,1820327937...1820328307&date=latest&sex=8&item=2&pay=1&measures=20100,20701")

@@

## Business counts

```{r}
# busi <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_141_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318,1870659585...1870659791,1870659801,1870659792...1870659800,1879048193...1879048573,1879048583,1879048574...1879048582&date=latest&industry=163577857...163577874&employment_sizeband=0&legal_status=0&measures=20100")

@@

```{r}
data <- list(lad, job.dens, pop, labour, earnings, busi)
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = "lad19cd") %>%
  #distinct(.keep_all = T)
  unique()
sapply(data, function(x) sum(is.na(x)))
```

@@

```{r correlations}
data.cor <- data %>%
  dplyr::select(LAD.up.am19, LAD.down.am19,  
                LAD.tests.am19, LAD.up.pm19, 
                LAD.down.pm19, LAD.tests.pm19, 
                LAD.up.am20, LAD.down.am20,
                LAD.tests.am20, LAD.up.pm20, 
                LAD.down.pm20, LAD.tests.pm20)

@@

data.cor <- data %>%
  dplyr::select(LAD.up.am20, LAD.up.pm20,
                LAD.tests.pm20,
                pop, pop16_64,
                managers, tech,
                skilled, earnings,
                total.busi)

m <- cor(data.cor, use = "pairwise.complete.obs")
corrplot(m, type="upper",method = "number", number.cex = .5, tl.cex = .75)
```

## Auxiliary regression, dep. var.: upload clusters

```{r upload regression, results= 'asis', message=F, error=F, warning=F}
# helpful sources: http://www.princeton.edu/~otorres/LogitR101.pdf
# https://thestatsgeek.com/2014/02/08/r-squared-in-logistic-regression/

aux <- multinom(cluster.up ~ LAD.up.am20 + LAD.up.pm20 + LAD.tests.pm20 + #  LAD.down.am20 + LAD.down.pm20
                   pop + pop16_64 + # how about young population share
                   managers + tech + skilled +
                   earnings + total.busi, data = data, 
                trace = F) # to avoid messages

stargazer(aux, type = "html",
          add.lines = list(c("McFadden's R squared",                    # use rownames as line name. in this case Weak insterumetns
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3))))

@@

aux <- multinom(cluster.up ~ pop + pop16_64 + # how about young population share
                  managers + tech + skilled +
                  earnings + total.busi, data = data, 
                trace = F) # to avoid messages

@@

stargazer(aux, type = "html",
          add.lines = list(c("McFadden's R squared",                    # use rownames as line name. in this case Weak insterumetns
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3))))

@@

## Auxiliary regression, dep. var.: download clusters

```{r download regression, results= 'asis', message=F, error=F, warning=F}
# helpful sources: http://www.princeton.edu/~otorres/LogitR101.pdf
# https://thestatsgeek.com/2014/02/08/r-squared-in-logistic-regression/

aux <- multinom(cluster.down ~ LAD.down.am20 + LAD.down.pm20 + LAD.tests.pm20 + #  LAD.down.am20 + LAD.down.pm20
                   pop + pop16_64 + # how about young population share
                   managers + tech + skilled +
                   earnings + total.busi, data = data, 
                trace = F) # to avoid messages

# McFadden's R squared
aux.null <- multinom(cluster.down ~ 1, data = data)
r2 <- 1-logLik(aux)/logLik(aux.null)

stargazer(aux, type = "html",
          add.lines = list(c("McFadden's R squared",                    # use rownames as line name. in this case Weak insterumetns
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3), round(r2[1],3), round(r2[1],3),
                              round(r2[1],3))))

>>>>

```{r settings, include = FALSE}
library(ggplot2)
library(tidyverse)
library(rprojroot)
library(lubridate)
library(knitr)
library(rprojroot)
library(nnet)
library(stargazer)
library(openair)
library(rgdal)
library(corrplot)
library(DescTools)
library(geosphere)
library(rgeos)
library(raster)

@@

# LAD codes and names
# get LA 
# this is the heaviest version:
# (i) directly from the web
#la <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/b6d2e15801de45328b760a4f55d74318_0.geojson?outSR={%22latestWkid%22:3857,%22wkid%22:102100}")#, layer="OGRGeoJSON")
# or, (ii) from the json I saved locally
# path.json <- paste(path, "./data/raw/Local_Authority_Districts_(April_2019)_Boundaries_UK_BFE.geojson", sep = "")
# la <- readOGR(path.json)#, layer="OGRGeoJSON")
# UK BGC: 
la <- readOGR("https://opendata.arcgis.com/datasets/0e07a8196454415eab18c40a54dfbbef_0.geojson")
codes <- la@data %>%
  dplyr::select(lad19cd, lad19nm) %>%
  rename(LAD19NM = lad19nm)

@@

ladDist$distMet <- apply(ladDist[,3:14], 1, min)
ladDist <- ladDist %>%
  dplyr::select(lad19cd, distLondon, distMet)

@@

## Job density 2018

```{r}

job.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_57_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318&date=latest&item=3&measures=20100")

@@

## Population 2018

@@

## Population density 2018

```{r}

# this is just population
pop.dens <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_31_1.data.csv?geography=1820327937...1820328318,1816133633...1816133848&date=latestMINUS1-latest&sex=7&age=0,22&measures=20100")

pop.dens <- pop.dens %>%
  dplyr::select(GEOGRAPHY_CODE, GEOGRAPHY_NAME, AGE_NAME, OBS_VALUE) %>%
  distinct(GEOGRAPHY_CODE, AGE_NAME, .keep_all = T) %>% 
  spread(AGE_NAME, OBS_VALUE) %>% 
  rename(pop16_64 = `Aged 16 - 64`) %>%
  rename(pop = `All ages`) %>%
  rename(lad19cd = GEOGRAPHY_CODE) %>%
  dplyr::select(-pop16_64)

crs(la)
la$area_sqkm <- area(la) / 1000000

pop.dens <- merge(pop.dens, la@data, by.x = "lad19cd", by.y = "LAD19CD", all.x = T)

pop.dens <- pop.dens %>%
  mutate(pop.dens = pop/area_sqkm) %>%
  dplyr::select(lad19cd,pop.dens)

# sapply(pop.dens, function(x) sum(is.na(x)))
# test <- pop.dens[is.na(pop.dens$area_sqkm),]
```

## N/S divide

```{r}

# It only includes only the south

ns <- read_csv("https://opendata.arcgis.com/datasets/3ba3daf9278f47daba0f561889c3521a_0.csv") %>%
  mutate(south = ifelse(RGN19NM=="London" | 
                        RGN19NM=="South West" |
                        RGN19NM=="East of England" |
                        RGN19NM=="South East", 1, 0)) %>%
  dplyr::select(LAD19CD, south) %>% 
  dplyr::filter(south==1) %>%
  rename(lad19cd = LAD19CD)
```

## Labour supply March 2020

```{r}

labour <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_17_5.data.csv?geography=1816133633...1816133837,1820327937...1820328307&date=latest&variable=18,1532...1540,290,720&measures=20599,21001,21002,21003")

@@

## Earnings 2019

```{r}

earnings <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_30_1.data.csv?geography=1946157348,2092957698,2013265929,1816133633...1816133837,1820327937...1820328307&date=latest&sex=8&item=2&pay=1&measures=20100,20701")

@@

## Business counts 2019

```{r}

# busi <- read_csv("https://www.nomisweb.co.uk/api/v01/dataset/NM_141_1.data.csv?geography=1816133633...1816133848,1820327937...1820328318,1870659585...1870659791,1870659801,1870659792...1870659800,1879048193...1879048573,1879048583,1879048574...1879048582&date=latest&industry=163577857...163577874&employment_sizeband=0&legal_status=0&measures=20100")

@@

```{r}
data <- list(lad, ladDist, job.dens, pop, pop.dens, labour, earnings, busi) # ns 
sapply(data, function(x) dim(x))
sapply(data, function(x) names(x))

# merge with reduce
data <- data %>% 
  reduce(inner_join, by = "lad19cd") %>%
  #distinct(.keep_all = T)
  unique()
sapply(data, function(x) sum(is.na(x)))

data <- merge(data, ns, all.x = T)
data$south <- ifelse(is.na(data$south),0, data$south)
```

@@

```{r correlations, results = 'asis'}
data.cor <- data %>%
  dplyr::select(LAD.up.am19, LAD.down.am19,  
                LAD.tests.am19, LAD.up.pm19, 
                LAD.down.pm19, LAD.tests.pm19, 
                LAD.up.am20, LAD.down.am20,
                LAD.tests.am20, LAD.up.pm20, 
                LAD.down.pm20, LAD.tests.pm20)

@@

data.cor <- data %>%
  dplyr::select(LAD.tests.am20,
                pop, pop16_64,
                managers, tech,
                skilled, earnings,
                total.busi)

data.cor <- data %>%
  dplyr::select(pop16_64, 
                pop.dens,
                job.dens2018, 
                distMet,
                distLondon,
                south,
                managers, 
                tech,
                skilled,
                prof,
                admin, 
                caring,
                plant,
                earnings,
                total.busi,
                LAD.tests.am20,
                pop) %>%
  mutate('LAD.tests.am20/pop' = LAD.tests.am20/pop) %>%
    mutate('total.busi/pop' = total.busi/pop) %>%
  dplyr::select(-LAD.tests.am20,-pop, -total.busi)

m <- cor(data.cor, use = "pairwise.complete.obs")
corrplot(m, type="upper",method = "number", number.cex = .5, tl.cex = .75)

stargazer(data.cor,
          type="html",
          #type="text", 
          summary = T,
          nobs = T)
```

@@

# convert dependent variable to factor
data$cluster.up <- as.factor(data$cluster.up)
data$cluster.up <- relevel(data$cluster.up, "5") # base category: cluster 5

aux <- multinom(as.factor(cluster.up) ~ pop16_64 + # how about young population share, pop +
                  pop.dens + job.dens2018 + # pop.dens is still insignificant when job.dens is excluded
                  distMet + distLondon + south +
                  managers + tech + skilled + prof + admin + caring + plant +
                  earnings + I(total.busi/pop) +
                  I(LAD.tests.am20/pop), data = data, 
                trace = F) # to avoid messages

@@

stargazer(aux, 
          type = "html",
          #type = "text",
          nobs = T,
          covariate.labels=c("pop. 16-64, 2018",
                             "pop. densuty, 2018",
                             "job density, 2018",
                             "distance to nearest met. area",
                             "distance to London",
                             "South of the UK",
                             "% of managerial jobs, 2020",
                             "% of tech jobs, 2020",
                             "% of skilled trade jobs, 2020",
                             "% of professional jobs, 2020",
                             "% of dministrative jobs, 2020",
                             "% of leisure jobs, 2020",
                             "% of machine operation jobs, 2020",
                             "earnings, 2019",
                             "n. business est. per hab., 2019",
                             "n. of AM broadband tests per hab., 2019"),
          add.lines = list(c("McFadden's R squared",
                             round(r2[1],3), round(r2[1],3), round(r2[1],3),
                             round(r2[1],3), round(r2[1],3), round(r2[1],3),
                             round(r2[1],3), round(r2[1],3), round(r2[1],3)),
                           (c("N",
                            nrow(data), nrow(data), nrow(data),
                            nrow(data), nrow(data), nrow(data),
                            nrow(data), nrow(data), nrow(data)))))

@@
